Estimating length of SARS-CoV-2 from doubly interval censored data with missed infections

### 2024-03-12 paper-surv-analysis.tex
rearrange bits of paper

The format requirements from JASA are at \url{https://www.tandfonline.com/journals/uasa20}.
The most important parts are reproduced here.
\begin{itemize}
    \item Should be written with the following elements in the following order: title page; author footnote; abstract; keywords; article text (table(s); figures); acknowledgments; appendices; references
    \item Should be no more than 35 pages, including the abstract, figures, tables, and references. Appendices, proofs, and other supporting material should be placed in a separate supplement file (anonymized for review)
    \item Should contain an unstructured abstract of 200 words.
    \item Should contain between 3 and 5 keywords. Read making your article more discoverable, including information on choosing a title and search engine optimization.
    \item JASA requires that all papers be formatted for 8 1/2 x 11-inch paper, one side only.
    \item 12 point, double-spaced font (which we define as 26 lines of text per page)
    \item the average JASA paper is about 30 manuscript pages (with the above page format and including all appendices, references, tables, and figures), and it is very uncommon for papers much longer than about 35 pages to appear in JASA. 
\end{itemize}

Notes / thoughts on current version
\begin{itemize}
    \item Some of the derivations can be moved to a supplement (cut 2--3 pages).
    \item Need to rearranage the ATACCC stuff --- add some details of what I did (thesis ch 4) and move stuff from prior info section into supplement.
    \item Probably don't need some of the figures (maybe 5--8)?
    \item Should the sensitivity analyses (figs 13/14 be in the main text)?
    \item The above changes make this about the right length.
\end{itemize}

Table of contents, cut for final version but useful for now to compare lengths of sections.

@@

@@

The CIS had a household-based design inviting all individuals aged 2 and over from selected households.
Once invited, an enrolment swab would be taken at the first visit followed by 4 further weekly visits (giving a total of 5 swabs on days 0, 7, 14, 21, 28 relative to enrolment) after which visits were monthly.
As well as a swab to test for SARS-CoV-2, participants were asked to complete a questionnaire, and some participants were asked to provide blood samples.
In reality, visits were often not on this precise schedule, and occasionally missed.
A full description of the study can be found in the study protocol~\autocite{cisProtocol}.

Positive tests are grouped into \emph{infection episodes} which are assumed to originate from the same infection (procedure described in \cref{sec:episodes}).
I base my analysis on the 4,800 episodes that were first detected between 10th Oct 2020 and 6th Dec 2020 inclusive with negatives bounding the start and end time of the episode.
That is, the individual tested negative before and after the episode.
Without a negative before the episode there is no lower bound on the episode's start time.
It could be arbitrarily long and there is little information on its length.
A negative test follows each episode unless the individual was lost to follow-up before their episode ended.
This is rare.
Episodes removed in this step are incorporated in the adjustment for undetected episodes.

@@

Undetected events where there is no information on these events is known as \emph{truncation}.
A situation that is particularly relevant is \emph{left truncation}, when only events lasting a minimum amount of time are observed.
There is a subtlety that makes the CIS different from standard left truncation.
In CIS, individuals in which undetected infections occurred are enrolled in the study.
Therefore, some data has been collected about them.
In standard left truncation, nothing is known about the missed events.
Here, I will incorporate the additional information into the analysis.
I refer to these events as ``undetected'', rather than truncated, to emphasise this difference.

@@

Let $\na = (n_{1a}, \dots, n_{\nsched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\nsched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
&p(\na \mid \ntot, \theta) \\
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{\nsched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{\nsched} \left(\frac{1}{\nsched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

The posterior density we are interested in is $p(\theta \mid \na)$.
It can be written as follows.
\begin{align}
&p(\theta \mid \na) \\
&\propto p(\theta) p(\na \mid \theta) \\
&= p(\theta) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) p(\na \mid \theta, \ntot) \\
&\propto p(\theta) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \right) \\
&= p(\theta) \left( \prod_{i \in \set{D}} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:full-posterior}.
\end{align}
The rest of this section derives expressions for each $p_{ia}$, $p_{u}$ and the sum.

@@

\subsubsection{Generative model for false negatives} \label{imperf-test:sec:simulate}

\todo[inline]{Probably the wrong place for this subsection, and can be edited for brevity a lot}

\subsubsection{Incorporation into the survival analysis framework}

@@

First, in \cref{imperf-test:sec:modifying-p_ia}, I modify $p_{ia}$, the probability of an episode with test schedule $i$ being admissible, to allow for the episode possibly being longer than observed.
Then, in \cref{imperf-test:sec:modifying-p_iu}, I modify $p_{iu}$, the probability of an episode with test schedule $i$ being undetected, to allow for additional episodes being missed.

\subsection{Modifying \texorpdfstring{$p_{ia}$}{pia}} \label{imperf-test:sec:modifying-p_ia}

@@

\subsection{Modifying \texorpdfstring{$p_{iu}$}{piu}} \label{imperf-test:sec:modifying-p_iu}

>>>>

\textbf{Notes / thoughts on current version}
\begin{itemize}
    \item Some of the derivations can be moved to a supplement (cut 2--3 pages).
    \item Need to rearranage the ATACCC stuff --- add some details of what I did (thesis ch 4) and move stuff from prior info section into supplement.
    \item Probably don't need some of the figures (maybe 5--8)?
    \item Should the sensitivity analyses (figs 13/14 be in the main text)?
    \item The above changes make this about the right length.
\end{itemize}

The \textbf{format requirements from JASA} are at \url{https://www.tandfonline.com/journals/uasa20}.
The most important parts are reproduced here.
\begin{itemize}
    \item Should be written with the following elements in the following order: title page; author footnote; abstract; keywords; article text (table(s); figures); acknowledgments; appendices; references
    \item Should be no more than 35 pages, including the abstract, figures, tables, and references. Appendices, proofs, and other supporting material should be placed in a separate supplement file (anonymized for review)
    \item Should contain an unstructured abstract of 200 words.
    \item Should contain between 3 and 5 keywords. Read making your article more discoverable, including information on choosing a title and search engine optimization.
    \item JASA requires that all papers be formatted for 8 1/2 x 11-inch paper, one side only.
    \item 12 point, double-spaced font (which we define as 26 lines of text per page)
    \item the average JASA paper is about 30 manuscript pages (with the above page format and including all appendices, references, tables, and figures), and it is very uncommon for papers much longer than about 35 pages to appear in JASA. 
\end{itemize}

@@

\todo[inline]{Cut table of contents for final version but useful for now to compare lengths of sections.}

@@

The CIS had a household-based design inviting all individuals aged 2 and over from selected households.
Once invited, an enrollment swab would be taken at the first visit followed by 4 further weekly visits (giving a total of 5 swabs on days 0, 7, 14, 21, 28 relative to enrollment) after which visits were monthly.
Swabs were tested for the presence of SARS-CoV-2 using RT-PCR.
In reality, visits were often not on this precise schedule, and occasionally missed.
A full description of the study can be found in the study protocol~\autocite{cisProtocol}.

Positive tests are grouped into \emph{infection episodes} which are assumed to originate from the same infection (using the procedure described in \cref{sec:episodes}).
I base my analysis on the 4,800 episodes that were first detected between 10th Oct 2020 and 6th Dec 2020 inclusive with negatives bounding the start and end time of the episode.
That is, the individual tested negative before and after the episode.
Without a negative before the episode there is no lower bound on the episode's start time.
It could be arbitrarily long and there is little information on its length.
A negative test follows each episode unless the individual was lost to follow-up before their episode ended.
This is rare.
Episodes removed in this step are incorporated in the adjustment for undetected episodes.

@@

Undetected events where there is no information on these events is known as \emph{truncation}.
% A situation that is particularly relevant is \emph{left truncation}, when only events lasting a minimum amount of time are observed.
There is a subtlety that makes the CIS different from standard left truncation.
In CIS, individuals in which undetected infections occurred are enrolled in the study.
Therefore, some data has been collected about them.
In standard left truncation, nothing is known about the missed events.
Here, I will incorporate the additional information into the analysis.
I refer to these events as ``undetected'', rather than truncated, to emphasise this difference.

@@

Let $\na = (n_{1a}, \dots, n_{\nsched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\nsched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
p(\na \mid \ntot, \theta)
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{\nsched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{\nsched} \left(\frac{1}{\nsched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

The posterior density we are interested in is $p(\theta \mid \na)$.
It can be written as follows.
\begin{align}
p(\theta \mid \na)
&\propto p(\theta) p(\na \mid \theta) \\
&= p(\theta) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) p(\na \mid \theta, \ntot) \\
&\propto p(\theta) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \right) \\
&= p(\theta) \left( \prod_{i \in \set{D}} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:full-posterior}.
\end{align}
The rest of this section derives expressions for each $p_{ia}$, $p_{u}$ and the sum.

@@

@@

\subsubsection{Modifying \texorpdfstring{$p_{ia}$}{pia}} \label{imperf-test:sec:modifying-p_ia}

@@

\subsubsection{Modifying \texorpdfstring{$p_{iu}$}{piu}} \label{imperf-test:sec:modifying-p_iu}
